---
title: "HW_week4"
author: "SHUYUAN JIA"
date: "10/3/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

## GitHub URL

*https://github.com/Shuyuan301/STAT433_Fall2021.git*

```{r}
library(nycflights13)
library(tidyverse)
```


## Introduction
We know that there are many factors may affect the delays of flights. In this assignment, I will mainly focus on *Time of Flight*, *weather* (including temperature, humidity, and visibility), *Airlines* and *Airport* factors.

The following are some patterns I found:

1. Morning flights have less delays than evening flights. (Time of Flight as a factor)

2. Arrival delays of flights are closely related to the weather conditions. Extreme temperatures will cause more delays. As relative humidity increases, there is a slight increase for arrival delays. As the visibility improves, the times for arrival delays decrease. (Temperature, Humidity, and Visibility as factors)

3. The airline for the flight is also an important factor considering arrival delays. Some airlines always have less delays than others, such as Alaska Airlines, Hawaiian Airlines; while some airlines usually have more delays than others. (Airlines as a factor)



## Finding 1

Firstly, I consider the question that what time of day can avoid delays as much as possible. I group by the hour of the flight, and compute the mean value of arrival delays. Then arrange the mean arrival delays ascending.

```{r}
mean_delay_time = flights %>% 
  group_by(hour) %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(mean_arr_delay)

mean_delay_time
```

Through the results above, we can see that morning flights(from 5-9) have less delays, while the evening flights seem have more delays. Therefore, the earlier the flight is scheduled, then the lower its expected delay. Since the delays will affect later flights, and morning flights have fewer previous flights.


To make my finding more clear and powerful, I plot the mean arrival delay vs the time of flight below.

```{r}
ggplot(mean_delay_time,aes(x = hour, y =mean_arr_delay)) +
  geom_line()+
  scale_x_continuous(labels = as.character(mean_delay_time$hour),
                     breaks = mean_delay_time$hour) +
  labs(x = "Time of Flight", y = "Mean Arrial Delay",
       title = "Mean Arrival Delay vs Time of Flight") +
  theme(plot.title = element_text(hjust = 0.5))

```

From this plot, we can see clearly that the best flight time to avoid delays is morning.



### Finding 2:

There is a common sense that weather conditions are easily affect the status of flights. Therefore, I am going to figure out how these weather factors affect the flight delays. I choose to use temperature, humidity, and visibility as factors and see how they play a big role in this case.

In order to do so, I also use the relative data which is *weather*, and then combine them together.

```{r}
newdf = flights %>%
  select(-year, -month,-day, -hour) %>% 
  left_join(weather, by = c("origin", "time_hour"))

```

After that, I group by temperature, and compute the mean value of arrival delays. Then arrange the mean arrival delays descending.
```{r}
temp_df = newdf %>% 
  group_by(temp) %>% 
  summarise(arr_delay_mean = mean(arr_delay, na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(desc(arr_delay_mean))

temp_df
```

Through the result above, we may have a basic idea of how temperature affects the delays of flights. Let's look at the plot to get more direct conclusion.

```{r}
ggplot(temp_df,aes(x = temp, y = arr_delay_mean))+ geom_point() +
  geom_smooth(method = 'loess',se = FALSE, formula = 'y~x')+
  labs(x = "Temperature in F",
       y = "Mean Arrival Delay",
       title = "Mean Arrival Delay vs Temperature")+
  theme(plot.title = element_text(hjust = 0.5))
```


Through the plot, we can see clearly that at extreme temperatures, there are more delays.




Next, let's see how humidity affects the delays. Group by humidity, and compute the mean value of arrival delays. Then arrange the mean arrival delays descending.

```{r}
humid_df = newdf %>% 
  group_by(humid) %>% 
  summarise(arr_delay_mean = mean(arr_delay, na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(desc(arr_delay_mean))

humid_df
```


```{r}
ggplot(humid_df,aes(x = humid, y = arr_delay_mean))+ geom_point() +
  geom_smooth(method = 'loess',se = FALSE, formula = 'y~x')+
  labs(x = "Relative Humidity",
       y = "Mean Arrival Delay",
       title = "Mean Arrival Delay vs Relative Humidity")+
  theme(plot.title = element_text(hjust = 0.5))
```

We can see that there is a slight increase trend for arrival delays as the relative humidity increases.



Last, let's see how visibility affects the delays. Similarly, group by visibility, and compute the mean value of arrival delays. Then arrange the mean arrival delays.

```{r}
visib_df = newdf %>% 
  group_by(visib) %>% 
  summarise(arr_delay_mean = mean(arr_delay, na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(arr_delay_mean)

visib_df
```


```{r}
ggplot(visib_df,aes(x = visib, y = arr_delay_mean))+ geom_point() +
  geom_smooth(method = 'loess',se = FALSE, formula = 'y~x')+
  labs(x = "Visibility in Miles",
       y = "Mean Arrival Delay",
       title = "Mean Arrival Delay vs Visibility")+
  theme(plot.title = element_text(hjust = 0.5))
```

We can see that as the visibility improves, the delay curve goes down which means the mean arrival delays decrease.


### Finding 3:


After thinking about weather, I am wondering is it possible that the airlines and airports also affect the delays of flights. Firstly, I try to explore the pattern between Arrival Delay and Airlines.

Here, we need to use relative data *airlines*, and then combine it with flights data.

```{r}
airline_delay <- flights %>%
  left_join(airlines,by="carrier") %>%
  group_by(name) %>%
  summarise(arr_delay_mean = mean(arr_delay,na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(arr_delay_mean)

airline_delay
```

Through the result above, we can see that some airlines have less delays than some other airlines. Let's use plot to check.


```{r}
new = left_join(airlines,flights,by="carrier")
new %>% 
  drop_na() %>% 
  ggplot(aes(x = name, y = arr_delay))+ geom_point() +
  theme(axis.text.x = element_text(angle = 90))+ 
  labs(x = "Airlines",
       y = "Arrival Delay",
       title = "Arrival Delay vs Airlines")+
  theme(plot.title = element_text(hjust = 0.5))

```

Instead of using mean value, I use arrival delays directly since the average value may be affected by some extreme values. (I am sorry that it could run a little bit slow.) We can see that SkyWest Airlines, Hawaiian Airlines and Alaska Airlines are some airlines that have less delays.


Then I try to figure out if there is a pattern between Arrival Delay and Origin.

```{r}
origin_delay <- flights %>%
  left_join(airlines,by="carrier") %>%
  group_by(origin) %>%
  summarise(arr_delay_mean = mean(arr_delay,na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(desc(arr_delay_mean))

origin_delay
```

Here, I compute the mean arrival delays for each origin. Through this result, we can say there is relatively high probability to have delays at EWR. Also, use the plot to check.

```{r}
flights %>% 
  drop_na() %>% 
  ggplot(aes(x = origin, y = arr_delay))+ geom_point() +
  labs(x = "Origin",
       y = "Arrival Delay",
       title = "Arrival Delay vs Origin")+
  theme(plot.title = element_text(hjust = 0.5))
```

Through the plot above, we can not see the differences between EWR and LGA clearly. Therefore, we can not conclude that the origin is a strong factor to affect the delays for flights. So, I did not put this into my findings part at the very beginning of this assignment.