---
title: "HW_week3"
author: "SHUYUAN JIA"
date: "9/26/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## GitHub URL

*https://github.com/Shuyuan301/STAT433_Fall2021.git*

```{r}
library(dplyr)
# install.packages("nycflights13")
library(nycflights13)
library(ggplot2)
```
## Question 1

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```
**Through the result above, we can see that there are 8255 flights have a missing dep_time. The departure delays(dep_delay), arrival time(arr_time) and arrival delays(arr_delay) are also missing for these rows. Therefore, these rows seem to be the canceled flights.**


## Question 2

Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

**Convert dep_time and sched_dep_time**

```{r}
dep_time_hours = ((flights$dep_time)%/% 100)
dep_time_mins = ((dep_time_hours*60) + ((flights$dep_time)%% 100))%% 1440
#dep_time_mins

sched_dep_time_hours = ((flights$sched_dep_time)%/% 100)
sched_dep_time_mins = ((sched_dep_time_hours*60) + ((flights$sched_dep_time)%% 100))%% 1440
#sched_dep_time_mins

flight_time = mutate(flights, dep_time_mins,sched_dep_time_mins)
```

**To get the departure times in the number of minutes since midnight, we need to divide dep_time by 100 to get the hours. Then multiply by 60 and add the remainder of dep_time divided by 100.**

**Note: When the time is 24:00, there is 1440 minutes since midnight, but it should be 0. We need to deal with it by using %%1440 to convert 24:00 be zero.**


**Show only relevant columns**
```{r}
select(flight_time, dep_time, dep_time_mins, sched_dep_time,sched_dep_time_mins)
```

## Question 3

Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

**Firstly, we are trying to figure out whether there is a pattern for canceled flights or not.**
```{r}
mutate(flights,cancel = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(canceled_num = sum(cancel),
    flights_num = n()) %>% 
  ggplot(aes(x = flights_num, y = canceled_num)) + geom_point()
```
**Through the plot above (flights_num against canceled_num), we can see that the number of flights canceled increases with the total number of flights.**

**Next, try to figure out whether there is a relationship between the proportion of flights canceled and the average departure delay.**
```{r}
mutate(flights, cancel=is.na(arr_delay) | is.na(dep_delay)) %>% 
  group_by(year,month,day) %>% 
  summarise(canceled_prop = mean(cancel),
            avg_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = avg_dep_delay, y = canceled_prop)) + geom_point() 

```


**Through the plot above, we can see clearly that there is a strong increasing relationship between the proportion of flights canceled and the average departure delay.**


